---
title: "Ammodytes dubius Assignment 5"
output: github_document
---



# Setup

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
source("setup.R")
```

# Load the Brickman data
 
```{r load_covar}
cfg = read_configuration(scientificname = "Ammodytes dubius",
                         version = "v1", 
                         path = data_path("models"))
db = brickman_database()
db = brickman_database()
present_conditions = read_brickman(db |> filter(scenario == "PRESENT", 
                                                interval == "mon"),
                       add = c("depth", "month")) |>
  select(all_of(cfg$keep_vars))
```

# Load the workflow

We read the model information we created.

```{r load_workflow}
model_fits = read_model_fit(filename = "Ammodytes_dubius-v1-model_fits")
model_fits
```


# Make a prediction

First we make a "nowcast" which is just a prediction of the current environmental conditions. 

## Nowcast

First make the prediction using `predict_stars()' to show the current probability of Northern Sand Lance in different areas of the Gulf of Maine.

```{r nowcast}
nowcast = predict_stars(model_fits, present_conditions)
nowcast
```
Now we can plot what is  called a "habitat suitability index" (hsi) map. 

```{r plot_nowcast_maxent, warning = FALSE}
plot_prediction(nowcast['default_btree'])
```

We can also plot a presence/absence labeled map.

```{r plot_class_labels, warning = FALSE}
pa_nowcast = threshold_prediction(nowcast, threshold = 0.4)
plot_prediction(pa_nowcast['default_btree'])
```

## Forecast

Now we forecast for RCP85 and 45 in 2075 and 2055. We plot this data/ predictions on a map of the Gulf of Maine

```{r load_2075_RCP85, warning = FALSE}
covars_rcp85_2075 = read_brickman(db |> filter(scenario == "RCP85", 
                                               year == 2075, 
                                               interval == "mon"),
                                  add = c("depth", "month")) |>
  select(all_of(cfg$keep_vars))
covars_rcp85_2055 = read_brickman(db |> filter(scenario == "RCP85", 
                                               year == 2055, 
                                               interval == "mon"),
                                  add = c("depth", "month")) |>
  select(all_of(cfg$keep_vars))
```

```{r forecast}
forecast_2075_RCP85 = predict_stars(model_fits, covars_rcp85_2075)
forecast_2075_RCP85
forecast_2055_RCP85 = predict_stars(model_fits, covars_rcp85_2055)
forecast_2055_RCP85
```

```{r plot_forecast, warning = FALSE}
plot_prediction(forecast_2075_RCP85['default_btree'])
plot_prediction(forecast_2055_RCP85['default_btree'])
```
Now for RCP45!
```{r load_2075_RCP45, warning = FALSE}
covars_rcp46_2075 = read_brickman(db |> filter(scenario == "RCP85", 
                                               year == 2075, 
                                               interval == "mon"),
                                  add = c("depth", "month")) |>
  select(all_of(cfg$keep_vars))
covars_rcp45_2055 = read_brickman(db |> filter(scenario == "RCP85", 
                                               year == 2055, 
                                               interval == "mon"),
                                  add = c("depth", "month")) |>
  select(all_of(cfg$keep_vars))
```

```{r forecast2}
forecast_2075_RCP45 = predict_stars(model_fits, covars_rcp85_2075)
forecast_2075_RCP45
forecast_2055_RCP45 = predict_stars(model_fits, covars_rcp85_2055)
forecast_2055_RCP45
```
```{r plot_forecast2, warning = FALSE}
plot_prediction(forecast_2075_RCP45['default_btree'])
plot_prediction(forecast_2055_RCP45['default_btree'])
```
## Save the predictions

Now we save our predictions.

```{r save_pred}
# make sure the output directory exists
path = make_path("predictions")

write_prediction(nowcast,
                 scientificname = cfg$scientificname,
                 version = cfg$version,
                 year = "CURRENT",
                 scenario = "CURRENT")
write_prediction(forecast_2075_RCP85,
                 scientificname = cfg$scientificname,
                 version = cfg$version,
                 year = "2075",
                 scenario = "RCP85")
write_prediction(forecast_2055_RCP85,
                 scientificname = cfg$scientificname,
                 version = cfg$version,
                 year = "2055",
                 scenario = "RCP85")
write_prediction(forecast_2075_RCP45,
                 scientificname = cfg$scientificname,
                 version = cfg$version,
                 year = "2075",
                 scenario = "RCP45")
write_prediction(forecast_2055_RCP45,
                 scientificname = cfg$scientificname,
                 version = cfg$version,
                 year = "2055",
                 scenario = "RCP45")
```

