---
title: "Assignment 4"
output: github_document
---
```{r setup}
```
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("/home/lrkerm28/ColbyForecasting/setup.R")
```

## Load Data

The first step is to load the necessary data we need in order to create models based on the obis Grey Seal data. 
```{r load_data}
cfg = read_configuration(scientificname = "Ammodytes dubius", version = "v1")
model_input = read_model_input(scientificname = "Ammodytes dubius", 
                               version = "v1",
                               log_me = c("depth", "Xbtm")) |>
  dplyr::mutate(month = month_as_number(.data$month)) |>
  select(all_of(c("class", cfg$keep)))
```
## Splitting the data into groups

Next, we want to split the data into two groups: data for training the model during creation and data for testing the model after we have created it. After that, we want to split the **training** data.
## Initial split into training and testing groups
```{r initial_split}
model_input_split = spatial_initial_split(model_input, 
                        prop = 1 / 5,     # 20% for testing
                        strategy = spatial_block_cv) # see ?spatial_block_cv
model_input_split
```
```{r initial_split_plot}
autoplot(model_input_split)
```

## Split the training group into regional folds

Next we take just the **training** data and split that into a set cross-validation folds and create mini/smaller data sets.
```{r cv_training}
tr_data = training(model_input_split)
cv_tr_data <- spatial_block_cv(tr_data,
  v = 5,     
  cellsize = grid_cellsize(model_input),
  offset = grid_offset(model_input) + 0.00001
)
autoplot(cv_tr_data)
```
## Build a recipe

Now, we make a recipe, which is a blueprint that guides the data handling and modeling process.
```{r recipe}
one_row_of_training_data = dplyr::slice(tr_data,1)
rec = recipe(one_row_of_training_data, formula = class ~ .)
rec
```
```{r recipe_summary}
summary(rec)
```
## Create a workflow
We make a workflow in order to put the data into containers for storing the pre-processing steps and model specifications. we are using four types of models.

```{r make_workflow}
wflow = workflow_set(
  
  preproc = list(default = rec), # not much happening in our preprocessor
  
  models = list(                 # but we have 4 models to add
    
      # very simple - nothing to tune
      glm = logistic_reg(
          mode = "classification") |>
        set_engine("glm"),
      
      # two knobs to tune
      rf = rand_forest(
          mtry = tune(),
          trees = tune(),
          mode = "classification") |>
        set_engine("ranger", 
                   importance = "impurity"),
      
      # so many things to tune!
      btree = boost_tree(
          mtry = tune(), 
          trees = tune(), 
          tree_depth = tune(), 
          learn_rate = tune(), 
          loss_reduction = tune(), 
          stop_iter = tune(),
          mode = "classification") |>
        set_engine("xgboost"),
    
      # just two again
      maxent = maxent(
          feature_classes = tune(),
          regularization_multiplier = tune(),
          mode = "classification") |>
        set_engine("maxnet")
  )
)
wflow
```
## Chosing metrics for measuring how well the models do

Now, we next chose which metrics to use when evaluating model performance. 
```{r metrics}
metrics = sdm_metric_set(yardstick::accuracy)
metrics
```
## Fit the models to the various recipes
#  Hyperparameters

These parameters help guide the optimal structure for solving the problem and are set for each model.
### Iterate to find various hyperparameter sets

we will use a function that will take each recipe-model combination and iterator over each of the folds while adjusting the parameters . The results of each iteration will be tracked so we can get a sense of the range and average three times.
```{r fit, warning = FALSE}
wflow <- wflow |>
  workflow_map("tune_grid",
    resamples = cv_tr_data, 
    grid = 3,
    metrics = metrics, 
    verbose = TRUE)
```
Let's plot it!
```{r plot_wflow}
autoplot(wflow)
```
### Choose the best set of hyperparameters for each model
```{r select_best}
model_fits = workflowset_selectomatic(wflow, model_input_split,
                                  filename = "Ammodytes_dubius-v1-model_fits",
                                  path = data_path("models"))
model_fits
```
## Exploring the collection of model fit results
### A table of metrics
```{r model_fit_metrics}
model_fit_metrics(model_fits)
```
### Confusion matrices and accuracy

We can plot confusion matrices and add accuracy.
```{r model_fit_confmat}
model_fit_confmat(model_fits)
```
### ROC/AUC
```{r model_fit_roc_auc}
model_fit_roc_auc(model_fits)
```
### Variable importance

Variable importance tells us about the contribution each covariate variable makes toward the whole within a certain model. NOTE: This is still not working for me...
```{r model_fit_vip}
model_fit_varimp_plot(model_fits)
```
## Exploring a single model fit result
```{r random_forest}
rf = model_fits |>
  filter(wflow_id == "default_rf")
rf
```
```{r rf_splits}
autoplot(rf$splits[[1]])
```
```{r rf_metrics}
rf$.metrics[[1]]
```

```{r rf_preds}
rf$.predictions[[1]]
```
```{r rf_workflow}
rf$.workflow[[1]]
```
### Partial dependence plot

This reflects the relative contribution of each variable influence over it's full range of values. 
```{r pd_plot}
model_fit_pdp(model_fits, wid = "default_btree", title = "Boosted Tree")
```
