---
title: "Assignment 3"
output: github_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
source("/home/lrkerm28/ColbyForecasting/setup.R")
#source("setup.R")
```

```{r load_obs_mask, include=TRUE}
coast = read_coastline()
obs = read_observations(scientificname = "Ammodytes dubius")
db = brickman_database() |>
  filter(scenario == "STATIC", var == "mask")
mask = read_brickman(db)
```
## Read secondary species

Download the data from obis on Ammodytes dubius, then it can be made into a variable.
```{r read_obis_data, include=TRUE}
sand_lance = fetch_obis(scientificname = "Ammodytes dubius")
```
## Filter New Data

We use "read_observations()" to filter the NAs and not needed information out of the data. We then give it a summary to check that all of the NAs have been filtered out.
```{r filter_out_NAs, include=TRUE}
sand_lance = read_observations(scientificname = "Ammodytes dubius")
summary(sand_lance)
```
## Making Background and Presence Data
With code provided from the C02_background doc, we can see/make the background and save it in a readable file for the next tasks.



```{r all_observations_count}
LON0 = -67
LAT0 = 46
all_counts = count(st_drop_geometry(obs), month) # counting is faster without spatial baggage
all_counts
```
```{r all_observations}
ggplot() +
  geom_sf(data = obs, alpha = 0.2, shape = "circle small", size = 1) +
  geom_sf(data = coast, col = "orange") +
  geom_text(data = all_counts,
            mapping = aes(x = LON0, 
                          y = LAT0, 
                          label = sprintf("n: %i", .data$n)),
                          size = 3) + 
  labs(x = "Longitude", y = "Latitude", title = "All observations") +
  facet_wrap(~month)
```
## Thin Data

```{r thin_observations}
thinned_obs = sapply(month.abb,
               function(mon){ 
                 thin_by_cell(obs |> filter(month == mon), mask)
               }, simplify = FALSE) |>
  dplyr::bind_rows() 

# another count
thinned_counts = count(st_drop_geometry(thinned_obs), month)

ggplot() +
  geom_sf(data = thinned_obs, 
          alpha = 0.2, 
          shape = "circle small", 
          size = 1) +
  geom_sf(data = coast, col = "orange") +
  geom_text(data = thinned_counts,
            mapping = aes(x = LON0, 
                          y = LAT0, 
                          label = sprintf("n: %i", .data$n)),
                          size = 3) + 
  labs(x = "Longitude", y = "Latitude", title = "Thinned observations") +
  facet_wrap(~month)
```
## Weighted sampling
```{r bias_map}
bias_map = rasterize_point_density(obs, mask) # <-- note the original observations

ggplot() +
  geom_stars(data = bias_map, aes(fill = count)) +
  scale_fill_viridis_b(na.value = "transparent") +
  geom_sf(data = coast, col = "orange") + 
  labs(x = "Longitude", y = "Latitude", title = "Bias map using all observations")
```
## Randomly sample background points
```{r n_background_per_month}
nback_avg = mean(all_counts$n) |>
  round()
nback_avg
```
```{r sample_background}
obsbkg = sapply(month.abb,
    function(mon){ 
      sample_background(thinned_obs |> filter(month == mon), # <- just this month
                       bias_map,
                       method = "bias",  # <-- it needs to know it's a bias map
                       return_pres = TRUE, # <-- give me the obs back, too
                       n = nback_avg) |>   # <-- how many points
        mutate(month = mon, .before = 1)
    }, simplify = FALSE) |>
  bind_rows() |>
  mutate(month = factor(month, levels = month.abb))
obsbkg 
```
## Last step to save it
```{r save_model_input}
write_model_input(obsbkg, scientificname = "Ammodytes dubius")
```
```{r read_model_input}
x = read_model_input(scientificname = "Ammodytes dubius")
```

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
